Jean Pockets and Gender\thanks{Code and data are available at: https://github.com/hari-lr/pockets-and-gender}

### 2024-04-24 paper/paper.qmd
update

Throughout history, women have faced numerous restrictions and limitations, often ingrained within societal norms and cultural practices. Among these limitations, the realm of fashion has served as both a reflection and an enforcer of gender-based inequalities. From ancient times to the modern era, clothing has been a means of expressing social status, cultural identity, and gender roles. According to design expert, Hannah Carlson, pocket inequality stems from the way clothes are designed and made, she explains that: "From a very early age, I think we sort of agree as a culture that womenswear, girlswear is meant to be pretty," she said. "And menswear, boyswear is meant to be utilitarian" [@carlson]. Men's garments historically incorporated functional pockets, allowing them to carry essentials such as money, keys, and tools, women's clothing has often lacked this practical feature. Instead, women's garments have frequently been designed with form-fitting silhouettes, delicate fabrics, and ornamental details, prioritizing aesthetics over utility.However, the 1880s, women began to be vocal about pocket inequality, as the "demand for the vote and the demand for pockets were made together" [@carlson].

@@

\newpage
# Results
A merged scatterplot was generated to compare pocket area and price for both men's and women's jeans. Each gender category was represented by a distinct color in the plot. Additionally, linear regression lines were overlaid for each gender group to observe potential differences in the relationship between pocket area and price (see @fig-price_per_pocket_gender).

@@

In summary, the results suggest that pocket area is not a significant predictor of price for either men's or women's jeans in the dataset analyzed. Other factors not considered in this analysis are likely to have a greater influence on the pricing of jeans for both genders.

@@

## First discussion point {#sec-first-point}
## Second discussion point
## Third discussion point

##Weaknesses and next steps

>>>>

Throughout history, women have faced numerous restrictions and limitations, often ingrained within societal norms and cultural practices. Among these limitations, the realm of fashion has served as both a reflection and an enforcer of gender-based inequalities. From ancient times to the modern era, clothing has been a means of expressing social status, cultural identity, and gender roles. According to design expert, Hannah Carlson, pocket inequality stems from the way clothes are designed and made, she explains that: "From a very early age, I think we sort of agree as a culture that womenswear, girlswear is meant to be pretty," she said. "And menswear, boyswear is meant to be utilitarian" [@carlson]. Men's garments historically incorporated functional pockets, allowing them to carry essentials such as money, keys, and tools, women's clothing has often lacked this practical feature. Instead, women's garments have frequently been designed with form-fitting silhouettes, delicate fabrics, and ornamental details, prioritizing aesthetics over utility. However, the 1880s, women began to be vocal about pocket inequality, as the "demand for the vote and the demand for pockets were made together" [@carlson].

@@

# Results
A merged scatterplot was generated to compare pocket area and price for both men's and women's jeans. Each gender category was represented by a distinct color in the plot. Additionally, linear regression lines were overlaid for each gender group to observe potential differences in the relationship between pocket area and price (see @fig-price_per_pocket_gender).

@@

In summary, the results suggest that pocket area is not a significant predictor of price for either men's or women's jeans in the dataset analyzed. Other factors not considered in this analysis are likely to have a greater influence on the pricing of jeans for both genders.
\newpage

@@

## First discussion point

as mentioned before

## Second discussion point
## Third discussion point
https://www.mydoh.ca/learn/blog/lifestyle/what-is-the-pink-tax-a-guide-for-teens/

## Weaknesses and next steps

The analysis has several limitations. Firstly, it narrowly focuses on the relationship between pocket area and price, neglecting other potential pricing determinants in the fashion industry. The dataset's limited scope, derived from measurements of specific brands and styles, may not be representative of the broader jeans market. Additionally, the regression analysis yielded statistically insignificant results for pocket area coefficients in both men's and women's jeans, indicating a weak relationship. Moreover, the unknown units of pocket area measurement limits accurate interpretation.

To address these limitations, future research could incorporate additional pricing determinants like brand reputation and market demand. Diversifying the dataset to include jeans from additional brands and price points would enhance representativeness. Qualitative research methods could provide deeper insights into consumer behavior, offering a more comprehensive understanding of purchasing decisions beyond pocket considerations. Considering cultural and societal factors influencing fashion trends would offer valuable context for interpreting the results. Moreover, it would be important to increment the number of jeans measured to improve the robustness and generalizability of the findings.

### 2024-04-24 paper/paper.tex
update

\title{Jean Pockets and Gender\thanks{Code and data are available at:
https://github.com/hari-lr/pockets-and-gender}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{subtitle}
\author{Hari Lee Robledo}
\date{April 23, 2024}

\begin{document}
\maketitle
\begin{abstract}
Fashion has long reflected and reinforced gender inequalities, with
women's clothing often lacking functional pockets found in men's
garments. A 2018 study by Jan Diehm \& Amber Thomas on The Pudding
revealed this discrepancy, highlighting inferior pocket functionality in
women's jeans. This paper aims to reproduce these findings and explore
if gender disparities extend to pricing. Using the same data, we analyze
the correlation between prices of men's and women's jeans. Results
suggest that women's jeans not only have smaller pockets but also tend
to be more expensive. This study sheds light on gender inequalities in
everyday products and their broader implications. {[}UPDATE according to
findingsThe findings revealed that pocket area alone may not be a strong
predictor of jeans prices for either gender, and there may be other
factors influencing the pricing of jeans.{]}
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, enhanced, boxrule=0pt, breakable]}{\end{tcolorbox}}\fi

@@

Throughout history, women have faced numerous restrictions and
limitations, often ingrained within societal norms and cultural
practices. Among these limitations, the realm of fashion has served as
both a reflection and an enforcer of gender-based inequalities. From
ancient times to the modern era, clothing has been a means of expressing
social status, cultural identity, and gender roles. According to design
expert, Hannah Carlson, pocket inequality stems from the way clothes are
designed and made, she explains that: ``From a very early age, I think
we sort of agree as a culture that womenswear, girlswear is meant to be
pretty,'' she said. ``And menswear, boyswear is meant to be
utilitarian'' (Alwahaidi 2018). Men's garments historically incorporated
functional pockets, allowing them to carry essentials such as money,
keys, and tools, women's clothing has often lacked this practical
feature. Instead, women's garments have frequently been designed with
form-fitting silhouettes, delicate fabrics, and ornamental details,
prioritizing aesthetics over utility.However, the 1880s, women began to
be vocal about pocket inequality, as the ``demand for the vote and the
demand for pockets were made together'' (Alwahaidi 2018).

@@

In this paper, the dataset used for analysis was obtained from
\texttt{measurementRectangles.json}. This dataset mirrors the
measurements found in \texttt{measurements.csv}, including brand, style,
gender specification (men or women), product name, brand size, waist
size, fabric composition, price, and height and width measurements of
jean pockets (see Table~\ref{tbl-raw_data}). Notably, the dataset
contains a ``pocketArea'' column representing the area of the polygon
generated from a pocket's measurements. However, it's important to note
that the units of measurement for this area are not explicitly specified
in the provided metadata. To determine the units, one must examine how
the measurements were collected and processed. Since the measurements
were processed using computer-generated calculations and intended for
web display, it is possible that the pocket area is denoted in square
centimeters (px²). However, without explicit confirmation from the data
documentation or creators of the dataset, absolute certainty regarding
the units of measurement remains elusive.

@@

To visually explore the relationship between pocket area and price for
both men's (see Figure~\ref{fig-price_per_pocket_men}) and women's jeans
(see Figure~\ref{fig-price_per_pocket_women}), scatter plots with
regression lines were created.

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-price_per_pocket_men-1.pdf}

@@

\caption{\label{fig-price_per_pocket_men}Jean Price vs Pocket Size
(Men)}

{\centering \includegraphics{paper_files/figure-pdf/fig-price_per_pocket_women-1.pdf}

@@

\caption{\label{fig-price_per_pocket_women}Jean Price vs Pocket Size
(Women)}

@@

To gain further insights and make predictions about {[}{]}, we will be
implementing a linear regression model.

@@

Linear regression seeks to determine the optimal parameters \(\beta_0\)
and \(\beta_1\) that minimize the discrepancies between predicted and
observed prices across varying pocket areas in the dataset. Through this
minimization process, the model constructs a line that most accurately
represents the association between pocket area and price. Consequently,
we can predict price values based on known pocket area measurements.
This regression analysis sheds light on how pocket area influences the
pricing of jeans for both men and women, offering valuable insights into
the pricing dynamics within each gender category.

@@

\hypertarget{tbl-men_model_summary}{}
\begin{table}

\caption{\label{tbl-men_model_summary}Summary of Linear Regression Model for Men's Jeans }Summary of Linear Regression Model for Men's Jeans}
\centering
\begin{tabular}[t]{l|c|c|c|c}
\hline
  & Estimate & Std. Error & t value & Pr(>|t|)\\
\hline
(Intercept) & 105.9002168 & 53.5254438 & 1.9785024 & 0.0551543\\
\hline
pocket\_area & -0.0022856 & 0.0049134 & -0.4651738 & 0.6444619\\
\hline
\end{tabular}
\end{table}

\hypertarget{tbl-women_model_summary}{}
\begin{table}

\caption{\label{tbl-women_model_summary}Summary of Linear Regression Model for Women's Jeans }Summary of Linear Regression Model for Women's Jeans}
\centering
\begin{tabular}[t]{l|c|c|c|c}
\hline
  & Estimate & Std. Error & t value & Pr(>|t|)\\
\hline
(Intercept) & 136.7159600 & 42.0324855 & 3.252626 & 0.0024018\\
\hline
pocket\_area & -0.0093445 & 0.0068483 & -1.364495 & 0.1804384\\
\hline
\end{tabular}
\end{table}

A scatterplot was generated to compare pocket area and price for both
men's and women's jeans. Each gender category was represented by a
distinct color in the plot. Additionally, linear regression lines were
overlaid for each gender group to observe potential differences in the
relationship between pocket area and price.

@@

\caption{\label{fig-price_per_pocket_gender}Jean Price vs.~Pocket Area
Size per Gender}

@@

After visualizing the linear regression for both men's and women's
jeans, I proceeded to analyze the results obtained from the regression
models:

The linear regression models for both men's and women's jeans revealed
interesting insights into the relationship between pocket area and
price.

For men's jeans, the coefficient estimate for pocket area (\(\beta_1\))
is -0.002286, indicating a very small and statistically insignificant
negative association between pocket area and price. This suggests that,
on average, changes in pocket area do not have a significant impact on
the price of men's jeans. The p-value associated with the pocket area
coefficient is 0.6445, which is much greater than the typical
significance level of 0.05. Additionally, the ( R\^{}2 ) value, which
measures the proportion of variance in the dependent variable explained
by the independent variable, is very low at 0.005662. Overall, these
results suggest that there is little evidence to suggest a meaningful
relationship between pocket area and price for men's jeans.

In contrast, for women's jeans, the coefficient estimate for pocket area
is -0.009345, also indicating a small negative association with price,
although it is slightly larger in magnitude compared to men's jeans.
However, like men's jeans, this coefficient is statistically
insignificant with a p-value of 0.1804. The ( R\^{}2 ) value for women's
jeans is slightly higher at 0.04671, indicating that pocket area
explains a slightly larger proportion of the variance in price compared
to men's jeans. Despite this, the overall relationship between pocket
area and price remains weak for women's jeans.

In summary, the results suggest that pocket area is not a significant
predictor of price for either men's or women's jeans in the dataset
analyzed. Other factors not considered in this analysis are likely to
have a greater influence on the pricing of jeans for both genders.

@@

\hypertarget{sec-first-point}{%
\subsection{First discussion point}\label{sec-first-point}}

@@

\#\#Weaknesses and next steps

\appendix

\hypertarget{appendix}{%
\section*{Appendix}\label{appendix}}
\addcontentsline{toc}{section}{Appendix}

>>>>

\title{Jean Pockets and Gender\thanks{Code and data are available at:
https://github.com/hari-lr/pockets-and-gender}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{subtitle}
\author{Hari Lee Robledo}
\date{April 24, 2024}

\begin{document}
\maketitle
\begin{abstract}
Fashion has long reflected and reinforced gender inequalities, with
women's clothing often lacking functional pockets found in men's
garments. A 2018 study by Jan Diehm \& Amber Thomas on The Pudding
revealed this discrepancy, highlighting inferior pocket functionality in
women's jeans. This paper aims to reproduce these findings and explore
if gender disparities extend to pricing. Using the same data, we analyze
the correlation between prices of men's and women's jeans. Results
suggest that women's jeans not only have smaller pockets but also tend
to be more expensive. This study sheds light on gender inequalities in
everyday products and their broader implications. {[}UPDATE according to
findingsThe findings revealed that pocket area alone may not be a strong
predictor of jeans prices for either gender, and there may be other
factors influencing the pricing of jeans.{]}
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, frame hidden, breakable, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

@@

Throughout history, women have faced numerous restrictions and
limitations, often ingrained within societal norms and cultural
practices. Among these limitations, the realm of fashion has served as
both a reflection and an enforcer of gender-based inequalities. From
ancient times to the modern era, clothing has been a means of expressing
social status, cultural identity, and gender roles. According to design
expert, Hannah Carlson, pocket inequality stems from the way clothes are
designed and made, she explains that: ``From a very early age, I think
we sort of agree as a culture that womenswear, girlswear is meant to be
pretty,'' she said. ``And menswear, boyswear is meant to be
utilitarian'' (Alwahaidi 2018). Men's garments historically incorporated
functional pockets, allowing them to carry essentials such as money,
keys, and tools, women's clothing has often lacked this practical
feature. Instead, women's garments have frequently been designed with
form-fitting silhouettes, delicate fabrics, and ornamental details,
prioritizing aesthetics over utility. However, the 1880s, women began to
be vocal about pocket inequality, as the ``demand for the vote and the
demand for pockets were made together'' (Alwahaidi 2018).

@@

In this paper, the dataset used for analysis was obtained from
\texttt{measurementRectangles.json}. This dataset mirrors the
measurements found in \texttt{measurements.csv}, including brand, style,
gender specification (men or women), product name, brand size, waist
size, fabric composition, price, and height and width measurements of
jean pockets (see Table~\ref{tbl-raw_data}). Notably, the dataset
contains a ``pocketArea'' column representing the area of the polygon
generated from a pocket's measurements. However, it's important to note
that the units of measurement for this area are not explicitly specified
in the provided metadata. To determine the units, one must examine how
the measurements were collected and processed. Since the measurements
were processed using computer-generated calculations and intended for
web display, it is possible that the pocket area is denoted in square
centimeters (px²). However, without explicit confirmation from the data
documentation or creators of the dataset, absolute certainty regarding
the units of measurement remains elusive. \newpage

@@

To visually explore the relationship between pocket area and price for
both men's and women's jeans scatter plots with regression lines were
created (see Figure 1 and 2).

@@

\begin{minipage}[t]{0.50\linewidth}

\raisebox{-\height}{

\includegraphics{paper_files/figure-pdf/fig-price_pocket-1.pdf}

@@

\caption{\label{fig-price_pocket-1}Jean Price vs Pocket Size}

@@

\end{minipage}%
%
\begin{minipage}[t]{0.50\linewidth}

\raisebox{-\height}{

\includegraphics{paper_files/figure-pdf/fig-price_pocket-2.pdf}

\caption{\label{fig-price_pocket-2}Jean Price vs Pocket Size}

@@

To gain further insights and make predictions about jeans pricing, a
linear regression model was implemented.

@@

The linear regression seeked to determine the optimal parameters
\(\beta_0\) and \(\beta_1\) that minimize the discrepancies between
predicted and observed prices across varying pocket areas in the
dataset. Through this minimization process, the model constructs a line
that most accurately represents the association between pocket area and
price. Consequently, we can predict price values based on known pocket
area measurements. This regression analysis sheds light on how pocket
area influences the pricing of jeans for both men and women, offering
valuable insights into the pricing dynamics within each gender category.

@@

A merged scatterplot was generated to compare pocket area and price for
both men's and women's jeans. Each gender category was represented by a
distinct color in the plot. Additionally, linear regression lines were
overlaid for each gender group to observe potential differences in the
relationship between pocket area and price (see
Figure~\ref{fig-price_per_pocket_gender}).

@@

\caption{\label{fig-price_per_pocket_gender}Jean Price vs.~Pocket Size
per Gender}

@@

After visualizing the linear regression for both men's and women's
jeans, the results obtained from the regression models were analyzed.

For men's jeans, the coefficient estimate for pocket area (\(\beta_1\))
is -0.002286, indicating a very small and statistically insignificant
negative association between pocket area and price. This suggests that,
on average, changes in pocket area do not have a significant impact on
the price of men's jeans. The p-value associated with the pocket area
coefficient is 0.6445, which is much greater than the typical
significance level of 0.05. Additionally, the ( R\^{}2 ) value, which
measures the proportion of variance in the dependent variable explained
by the independent variable, is very low at 0.005662. Overall, these
results suggest that there is little evidence to suggest a meaningful
relationship between pocket area and price for men's jeans (see
Table~\ref{tbl-men_lm}).

In contrast, for women's jeans, the coefficient estimate for pocket area
is -0.009345, also indicating a small negative association with price,
although it is slightly larger in magnitude compared to men's jeans.
However, like men's jeans, this coefficient is statistically
insignificant with a p-value of 0.1804. The ( R\^{}2 ) value for women's
jeans is slightly higher at 0.04671, indicating that pocket area
explains a slightly larger proportion of the variance in price compared
to men's jeans. Despite this, the overall relationship between pocket
area and price remains weak for women's jeans (see
Table~\ref{tbl-women_lm}).

In summary, the results suggest that pocket area is not a significant
predictor of price for either men's or women's jeans in the dataset
analyzed. Other factors not considered in this analysis are likely to
have a greater influence on the pricing of jeans for both genders.
\newpage

\hypertarget{tbl-men_lm}{}
\begin{table}
\caption{\label{tbl-men_lm}Summary of Linear Regression Model for Men's Jeans }\tabularnewline

\centering
\begin{tabular}{l|c|c|c|c}
\hline
  & Estimate & Std. Error & t value & Pr(>|t|)\\
\hline
(Intercept) & 105.9002168 & 53.5254438 & 1.9785024 & 0.0551543\\
\hline
pocket\_area & -0.0022856 & 0.0049134 & -0.4651738 & 0.6444619\\
\hline
\end{tabular}
\end{table}

\hypertarget{tbl-women_lm}{}
\begin{table}
\caption{\label{tbl-women_lm}Summary of Linear Regression Model for Women's Jeans }\tabularnewline

\centering
\begin{tabular}{l|c|c|c|c}
\hline
  & Estimate & Std. Error & t value & Pr(>|t|)\\
\hline
(Intercept) & 136.7159600 & 42.0324855 & 3.252626 & 0.0024018\\
\hline
pocket\_area & -0.0093445 & 0.0068483 & -1.364495 & 0.1804384\\
\hline
\end{tabular}
\end{table}

@@

\hypertarget{first-discussion-point}{%
\subsection{First discussion point}\label{first-discussion-point}}

as mentioned before

@@

https://www.mydoh.ca/learn/blog/lifestyle/what-is-the-pink-tax-a-guide-for-teens/

\hypertarget{weaknesses-and-next-steps}{%
\subsection{Weaknesses and next steps}\label{weaknesses-and-next-steps}}

The analysis has several limitations. Firstly, it narrowly focuses on
the relationship between pocket area and price, neglecting other
potential pricing determinants in the fashion industry. The dataset's
limited scope, derived from measurements of specific brands and styles,
may not be representative of the broader jeans market. Additionally, the
regression analysis yielded statistically insignificant results for
pocket area coefficients in both men's and women's jeans, indicating a
weak relationship. Moreover, the unknown units of pocket area
measurement limits accurate interpretation.

To address these limitations, future research could incorporate
additional pricing determinants like brand reputation and market demand.
Diversifying the dataset to include jeans from additional brands and
price points would enhance representativeness. Qualitative research
methods could provide deeper insights into consumer behavior, offering a
more comprehensive understanding of purchasing decisions beyond pocket
considerations. Considering cultural and societal factors influencing
fashion trends would offer valuable context for interpreting the
results. Moreover, it would be important to increment the number of
jeans measured to improve the robustness and generalizability of the
findings.
